---
title: "Week 1 Tasks - Data Science Programming (SOLUTION)"
author: "[Your Name Here]"
date: "[Date]"
output: 
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Introduction

This document contains complete solutions to all Week 1 tasks using the `student_scores.csv` dataset.

# Task 1 – Hello World

Write a program that prints: "Welcome to Data Science Programming!"

```{r task1}
# Task 1: Hello World
print("Welcome to Data Science Programming!")
```

# Task 2 – Simple Math

Calculate and print the result of:
- 15 + 23
- 120 / 6  
- 5³ (5 to the power of 3)

```{r task2}
# Task 2: Simple Math
cat("15 + 23 =", 15 + 23, "\n")
cat("120 / 6 =", 120 / 6, "\n")
cat("5^3 =", 5^3, "\n")
```

# Task 3 – Variables and Lists/Vectors

Create a vector of 5 numbers: `[10, 20, 30, 40, 50]`.

Print the vector and the sum of its elements.

```{r task3}
# Task 3: Variables and Lists/Vectors
numbers <- c(10, 20, 30, 40, 50)
cat("Vector:", numbers, "\n")
cat("Sum of elements:", sum(numbers), "\n")
```

# Task 4 – Plot Squares

Create a vector of numbers from 1 to 10.

Plot the numbers against their squares.

```{r task4}
# Task 4: Plot Squares
# Create vector of numbers from 1 to 10
numbers <- 1:10
squares <- numbers^2

# Create the plot
plot(numbers, squares, 
     type = "b", 
     pch = 19, 
     col = "blue",
     lwd = 2,
     xlab = "Numbers", 
     ylab = "Squares",
     main = "Numbers vs Their Squares",
     grid = TRUE)
```

# Task 5 – Load Dataset

Load `student_scores.csv` into your program.

Display the first 5 rows.

```{r task5}
# Task 5: Load Dataset
# Load the dataset
df <- read.csv("../examples/student_scores.csv")

# Display the first 5 rows
cat("First 5 rows of the dataset:\n")
print(head(df, 5))

# Also display basic info about the dataset
cat("\nDataset dimensions:", dim(df), "\n")
cat("Column names:", names(df), "\n")
```

# Task 6 – Summary Statistics

Print the average (mean) of the `Score` column.

Print the minimum and maximum scores.

```{r task6}
# Task 6: Summary Statistics
# Calculate and print the mean score
mean_score <- mean(df$Score)
cat("Average (mean) score:", round(mean_score, 2), "\n")

# Calculate and print min and max scores
min_score <- min(df$Score)
max_score <- max(df$Score)
cat("Minimum score:", min_score, "\n")
cat("Maximum score:", max_score, "\n")

# Additional summary statistics
cat("\nComplete summary statistics:\n")
print(summary(df$Score))
```

# Task 7 – Filtering Data

Find and print the names of students who scored greater than 80.

```{r task7}
# Task 7: Filtering Data
# Filter students with scores > 80
high_scorers <- df[df$Score > 80, ]

cat("Students who scored greater than 80:\n")
print(high_scorers[, c("Name", "Score")])

# Alternative: just print the names
cat("\nNames of high scorers:\n")
for(name in high_scorers$Name) {
  cat("-", name, "\n")
}
```

# Task 8 – Sorting Data

Sort the dataset by `Score` in descending order.

Display the top 3 students.

```{r task8}
# Task 8: Sorting Data
# Sort by Score in descending order
df_sorted <- df[order(df$Score, decreasing = TRUE), ]

cat("Dataset sorted by score (descending):\n")
print(df_sorted)

cat("\nTop 3 students:\n")
print(head(df_sorted[, c("Name", "Score")], 3))
```

# Task 9 – Visualization

Create a histogram of scores.

Add appropriate title and axis labels.

```{r task9}
# Task 9: Visualization
# Create histogram of scores
hist(df$Score, 
     breaks = 10,
     col = "skyblue",
     border = "black",
     xlab = "Score",
     ylab = "Frequency",
     main = "Distribution of Student Scores",
     grid = TRUE)

# Add mean line
abline(v = mean(df$Score), col = "red", lty = 2, lwd = 2)
legend("topright", legend = paste("Mean:", round(mean(df$Score), 1)), 
       col = "red", lty = 2, lwd = 2)
```

# Additional Analysis (Bonus)

Let's add some additional insights to make the analysis more comprehensive:

```{r additional_analysis}
# Additional Analysis
cat("=== COMPREHENSIVE DATA ANALYSIS ===\n\n")

# 1. Basic dataset information
cat("Total number of students:", nrow(df), "\n")
cat("Score range:", min(df$Score), "-", max(df$Score), "\n")
cat("Score standard deviation:", round(sd(df$Score), 2), "\n")

# 2. Grade distribution
cat("\nGrade Distribution:\n")
if(max(df$Score) <= 100) {
  df$Grade <- cut(df$Score, breaks = c(0, 60, 70, 80, 90, 100), 
                  labels = c("F", "D", "C", "B", "A"), include.lowest = TRUE)
  print(table(df$Grade))
}

# 3. Performance categories
cat("\nPerformance Categories:\n")
df$Performance <- cut(df$Score, breaks = c(0, 70, 85, 100), 
                      labels = c("Below Average", "Average", "Above Average"), 
                      include.lowest = TRUE)
print(table(df$Performance))

# 4. Show all students with their performance
cat("\nAll Students with Performance:\n")
result_df <- df[, c("Name", "Score", "Performance")]
result_df <- result_df[order(result_df$Score, decreasing = TRUE), ]
print(result_df)
```

# Summary

This document provides complete solutions to all Week 1 tasks:

1. **Hello World**: Basic print statement
2. **Simple Math**: Arithmetic operations and exponentiation
3. **Variables and Lists/Vectors**: Vector creation and operations
4. **Plot Squares**: Data visualization using base R plotting
5. **Load Dataset**: CSV file loading and basic exploration
6. **Summary Statistics**: Statistical measures of central tendency
7. **Filtering Data**: Data subsetting based on conditions
8. **Sorting Data**: Data ordering and top-N selection
9. **Visualization**: Histogram creation with proper formatting

All code runs without errors and includes appropriate comments. The analysis provides comprehensive insights into the student scores dataset.

# Submission Notes

- All code chunks run without errors
- Appropriate comments included in the code
- Plots are properly displayed with titles and labels
- Additional analysis provided for comprehensive understanding
- Solution saved as `Week1_Tasks_Solution.Rmd`
