---
title: "Week 3 Tasks — Data Preparation, EDA, and Intro Modeling (R)"
author: "Your Name"
output:
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(broom)
theme_set(theme_minimal())
```

# Instructions

- Use this template to complete Week 3 tasks. Replace placeholders with your work.
- Ensure the document can knit top-to-bottom without errors.
- Add short captions/annotations below each plot and metric output.

# Task 1 — Load Data and Inspect

```{r load-data}
# TODO: Set path and read your dataset
# Example:
# data <- read_csv("path/to/your.csv")
# head(data, 10)
# dim(data)
```

Briefly describe the dataset, its purpose, and key variables.

# Task 2 — Data Types, Summary Stats, and Missingness

```{r types-summary-missing}
# Glimpse types
# glimpse(data)

# Summary statistics (numeric and categorical)
# summary(data)

# Missingness per column
# tibble(col = names(data), n_missing = colSums(is.na(data))) %>%
#   mutate(p_missing = n_missing / nrow(data)) %>%
#   arrange(desc(p_missing))
```

# Task 3 — Data Cleaning

```{r cleaning}
# Examples:
# data <- distinct(data)
# names(data) <- names(data) %>%
#   tolower() %>% str_replace_all(" ", "_") %>% str_trim()
# data <- data %>% mutate(num_col = if_else(is.na(num_col), median(num_col, na.rm = TRUE), num_col))
```

# Task 4 — Exploratory Data Analysis (EDA)

```{r eda}
# Example univariate
# data %>% ggplot(aes(numeric_col)) + geom_histogram(binwidth = 5, fill = "#2c7fb8") +
#   labs(title = "Distribution of numeric_col", x = "numeric_col", y = "Count")

# Example bivariate
# ggplot(data, aes(x = xvar, y = yvar)) +
#   geom_point(alpha = 0.6) +
#   geom_smooth(method = "lm", se = FALSE, color = "#fc8d59") +
#   labs(title = "xvar vs yvar")
```

Write 3–5 short observations from EDA here.

# Task 5 — Data Visualization

```{r viz}
# Create at least 2 clear plots with captions below
# data %>% count(category) %>%
#   ggplot(aes(x = fct_reorder(category, n), y = n)) + geom_col(fill = "#7fcdbb") +
#   coord_flip() + labs(title = "Counts by category", x = "Category", y = "Count")
```

Key takeaway:



