---
title: "Week 5 Tasks — Data Visualization in R"
author: "Your Name"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(lubridate)
theme_set(theme_minimal())
```

# Instructions

- Use this template to complete Week 5 tasks. Replace placeholders with your work.
- Ensure the document can knit top-to-bottom without errors.
- Add short captions/annotations below each plot.

# Task 1 — Setup and Data Loading

```{r load-data}
# TODO: Set path and read your dataset
# Example:
# data <- read_csv("path/to/your.csv")
# head(data, 10)
```

Briefly describe the dataset and variables here.

# Task 2 — Univariate Visualizations

```{r univariate-numeric}
# Example histogram/density for a numeric variable
# ggplot(data, aes(numeric_col)) +
#   geom_histogram(binwidth = 5, fill = "#2c7fb8") +
#   labs(title = "Distribution of numeric_col", x = "numeric_col", y = "Count")
```

Key takeaway:

```{r univariate-categorical}
# Example bar plot for a categorical variable
# data %>% count(category) %>%
#   ggplot(aes(x = fct_reorder(category, n), y = n)) +
#   geom_col(fill = "#7fcdbb") + coord_flip() +
#   labs(title = "Counts by category", x = "Category", y = "Count")
```

Key takeaway:

# Task 3 — Bivariate Visualizations

```{r bivariate-scatter}
# Example scatter with smooth
# ggplot(data, aes(x = xvar, y = yvar)) +
#   geom_point(alpha = 0.6) +
#   geom_smooth(method = "lm", se = FALSE, color = "#fc8d59") +
#   labs(title = "xvar vs yvar")
```

Interpretation:

```{r bivariate-box}
# Example box/violin by category
# ggplot(data, aes(x = category, y = numeric_col)) +
#   geom_violin(fill = "#c7e9b4", alpha = 0.8) +
#   geom_boxplot(width = 0.15, outlier.alpha = 0.2) +
#   labs(title = "numeric_col by category")
```

Interpretation:

# Task 4 — Multivariate Visualizations

```{r multivariate-facets}
# Example facets
# ggplot(data, aes(x = xvar, y = yvar, color = group)) +
#   geom_point(alpha = 0.6) +
#   facet_wrap(~ facet_var) +
#   labs(title = "Relationship by facet_var")
```

Short note:

# Task 5 — Temporal or Composition Analysis

```{r temporal-line}
# Example time series (if applicable)
# ggplot(data, aes(x = date_col, y = value)) +
#   geom_line(color = "#2b8cbe") +
#   labs(title = "Value over time", x = "Date", y = "Value")
```

Alternative composition example if no time variable is present:

```{r composition}
# Example stacked percentage bar
# data %>% count(group, category) %>%
#   group_by(group) %>% mutate(p = n/sum(n)) %>% ungroup() %>%
#   ggplot(aes(x = group, y = p, fill = category)) +
#   geom_col() + scale_y_continuous(labels = scales::percent) +
#   labs(title = "Composition by group", y = "Percent", x = "Group")
```

# Task 6 — Geospatial (Bonus)

```{r geospatial, eval=FALSE}
# Optional: Use sf/leaflet/ggplot2 for maps if you have location data
# library(sf)
# library(leaflet)
```

# Task 7 — Aesthetics and Clarity

List the style choices you made (titles, labels, themes, color choices).

# Task 8 — Narrative and Reproducibility

Write 5–10 sentences connecting the plots into a cohesive story and add final insights.

